\documentclass[defaultstyle,11pt]{thesis}

\usepackage{amssymb}		% to get all AMS symbols
\usepackage{graphicx}		% to insert figures

\usepackage{rotating}
\usepackage{graphics}
\usepackage{colortbl}
\usepackage{color}
%\usepackage{hyperref}
\usepackage{subfig}
\usepackage{amsmath}

%%%%%%%%%%%%   All the preamble material:   %%%%%%%%%%%%

%\title{Devices for Embodied Fabrication}
\title{Embodied Fabrication:\\ Body-Centric Devices for Novice Designers}

\author{Benjamin A.}{Leduc-Mills}

\otherdegrees{B.A., University of Santa Cruz, 2003 \\
	      M.P.S., New York University, 2008 \\
	      M.S., University of Colorado, 2013}

\degree{Doctor of Philosophy}		%  #1 {long descr.}
	{Ph.D., Computer Science}		%  #2 {short descr.}

\dept{Department of}			%  #1 {designation}
	{Computer Science}		%  #2 {name}

\advisor{Prof.}				%  #1 {title}
	{Michael Eisenberg}			%  #2 {name}

\reader{Clayton Lewis}		%  2nd person to sign thesis
\readerThree{Tom Yeh}		%  3rd person to sign thesis

\abstract{  \OnePageChapter	% because it is very short
We present a class of devices under the umbrella moniker of ``embodied
fabrication''. These devices and the development of the term ``embodied
fabrication'' is rooted not only in computer science, but also in cognitive
science, children's educational theory, digital fabrication technology, and the
convergence of these strands present in the do-it-yourself community known as
the``maker movement''. As such, we operate under a certain set of premises that
guide and direct this work. First, that embodied cognition - which places the
body at the center of our cognitive operations - provides a framework in which
to ground our decisions to design physical peripheral devices as opposed to
purely screen-based software. Second, that a strong line of research supports
providing children with tangible, ``manipulative'' objects to learn with. Third,
that digital fabrication technologies - 3D printing in particular - provide a
wonderful new opportunity for children and novice designers in general to make,
play, and explore creatively - and that the current design options for 3D
printers are not suited to meaningful design and creation of objects by
non-expert 3D modelers. Finally, that by continuing in the best traditions of
body-centric interaction design for children, a suite of devices can be created
to provide an educationally and technically rich environment to connect kids to
the creative potential of 3D printing. We unpack these ideas more in the
introduction, followed by a discussion of all three of the prototype devices and
the companion software, a chapter on related work, a chapter discussing the
three user studies we performed evaluating these devices, and finally we present
a vision of the future of this work and of embodied fabrication devices as a
whole.
	}

\dedication[Dedication]{	% NEVER use \OnePageChapter here.
	To all the killas and the hundred dolla billas.
	}

\acknowledgements{ \OnePageChapter % *MUST* BE ONLY ONE PAGE! 

I was incredibly fortunate to have found the supporting cast that I had during
my time at Colorado. My advisor, Michael Eisenberg, is one of the true gems of
humanity. An unbelievable scholar, gifted writer, and a wonderfully supportive
and warm person. Over our many hours of discussion he taught me how to love and
appreciate the world where children's crafts, cognitive science, and
computation come together. My father, for urging me to visit Mike as I was
applying to grad school. You always let me know how proud you were.
You and I are more similar than I ever would have imagined as a child. I see
that as a wonderful thing. My mother, for her unwavering faith that I would be
excellent. You were much better at convincing me than I was at convincing
myself. Your visits were always a chance to recharge my batteries, feel loved,
and be fed extremely well.
Monika, for taking the risk of leaving your world behind to join me here in
Colorado, and for supporting me so well during a very stressful time. You're the
best cheerleader a guy could ask for. Many others contributed to this work,
academically and socially. I owe an enormous debt of gratitude to Nathan Siedle,
whose generosity allowed my work to reach its completion. My committee members
were all amazing. Ann Eisenberg kept the lab from self-destruction more times
than I can count. My CTG and CU compatriots for welcoming me to Colorado and
convincing me to stay. Lindsay Levkoff Diamond, for being an incredibly
understanding and flexible boss during my studies. The DOE and all my friends at
SparkFun Electronics - Toni, Amanda, Dave, Brennen, Jordan, (etc.) without you,
I don't know if I would have made it. And finally, Fran, Anna, Adrian, Victor,
Galen, and the amazing kids at Gold Crown for their support.
Thank you all.}

% \IRBprotocol{E927F29.001X}	% optional!


%\ToCisShort	% use this only for 1-page Table of Contents

%\LoFisShort	% use this only for 1-page Table of Figures
% \emptyLoF	% use this if there is no List of Figures

%\LoTisShort	% use this only for 1-page Table of Tables
% \emptyLoT	% use this if there is no List of Tables

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%       BEGIN DOCUMENT...         %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\begin{document}

% \tableofcontents
% \listofigures
% \listoftables

\input{macros.tex}

\input{1_introduction.tex}
\input{2_prototypeSystems.tex}
\input{3_RelatedWork.tex}
\input{4_Evaluation.tex}
\input{5_vision.tex}
\input{6_future_work.tex}
%%%%%%%%%   then the Bibliography, if any   %%%%%%%%%
\bibliographystyle{plain}	% or "siam", or "alpha", etc.
\nocite{*}		% list all refs in database, cited or not
\bibliography{refs}		% Bib database in "refs.bib"

%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
\appendix
\input {appendixA.tex}
\input {appendixB.tex}


\end{document}

